---
title: "Bulik_QC"
author: "Alicia Walker"
date: '2022-06-28'
output: html_document
---

# 1. Summary

```{bash}
#!/bin/bash
#
#SBATCH --job-name=summary
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

# Bulik
plink2 --bfile $DATA --missing  --out Summary
plink2 --bfile $DATA --freq     --out Summary
plink2 --bfile $DATA --hardy    --out Summary
```

# 2. Create Plots
export R_LIBS=/nfs/home/aliwal/rlib_plinkqc
```{r}
library(ggplot2)
library(ggpubr)
library(ggsignif)


vmiss <- read.table("Summary.vmiss", col.names = c("CHROM", "ID", "MISSING_CT", "OBS_CT", "F_MISS"))
smiss <- read.table("Summary.smiss", col.names=c("FID", "IID", "MISSING_CT", "OBS_CT", "F_MISS"))
freq <- read.table("Summary.afreq", col.names=c("CHROM", "ID", "REF", "ALT", "ALT_FREQS", "OBS_CT"))
hwe <- read.table("Summary.hardy", col.names=c("CHROM", "ID", "A1", "AX", "HOM_A1_CT", "HET_A1_CT", "TWO_AX_CT", "O(HET_A1)", "E(HET_A1)", "P"))

png("Bulik_Summary.png", width=18, height=18, units='in', bg="white", res=150, type=c("cairo"))
par(mfrow=c(2,2))
hist(1-vmiss$F_MISS, breaks="sturges",main="SNPs",col="tan", xlab="Genotyping Rate", ylab="Number of SNPs", labels=T, cex.lab = 2, cex.axis =2, cex.main =2)
hist(1-smiss$F_MISS, breaks="sturges", main="Individuals", col="tan", xlab="Genotyping Rate", ylab="Number of Individuals", cex.lab = 2, cex.axis =2, cex.main =2)
hist(hwe$P, breaks="sturges", main="HWE P-Value", col="tan", xlab="HWE P-value", ylab="Number of SNPs", cex.lab = 2, cex.axis =2, cex.main =2)
hist(freq$ALT_FREQS, breaks="sturges", main="MAF", col="tan", xlab="MAF", ylab="Number of SNPs", cex.lab = 2, cex.axis =2, cex.main =2)
dev.off()
```


# 3. SNP Missingness
Remove the SNPs with high missing rate (> 10%, which reflects poor performance of the SNP probe).
```{bash}

#!/bin/bash
#SBATCH --job-name=geno-missingness
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink2 --bfile $DATA --geno 0.1 --make-bed --out ${DATA}_cleaned1

```

grep removed *_cleaned1.log | grep missing

--geno: 9465 variants removed due to missing genotype data.

# 4.  Remove SNPS that violate HWE
HWE test threshold is set at 0.000001.
```{bash}
#!/bin/bash
#SBATCH --job-name=hwe
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink2 --bfile ${DATA}_cleaned1 --hwe 0.000001 --make-bed --out ${DATA}_cleaned2
```

grep removed *_cleaned2.log | grep hwe

--hwe:  7657 variants removed due to Hardy-Weinberg exact test (founders only).

# 5. Remove rare variants
MAF threshold is set at 0.01. The SNPs outside of chromosomes 1-23 are also removed in this step.
```{bash}
#!/bin/bash
#SBATCH --job-name=maf
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink2 --bfile ${DATA}_cleaned2 --maf 0.01  --chr 1-23 --make-bed --out ${DATA}_cleaned3
```

grep removed *_cleaned3.log | grep ' allele frequency threshold'
108860 variants removed due to allele frequency threshold(s)

# 6. Remove individuals with missing genotyping data

Remove individuals with high missing rate (> 10%, which reflects poor DNA quality)
```{bash}
#!/bin/bash
#SBATCH --job-name=missingness
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink2 --bfile ${DATA}_cleaned3 --mind 0.1 --make-bed --out ${DATA}_cleaned4
```

grep removed *_cleaned4.log | grep missing

125 samples removed due to missing genotype data (--mind).


# 7. Add Gender to .fam file

First add sex to the .fam file
```{r}
bulik <- read.csv("/nfs/home/aliwal/ANGI/data/sample_sheets/ANGI_Bulik_190430_sample.tsv", sep = "\t")
bulik <- bulik[,c("chip_well_barcode", "reported_gender")]
fam <- read.table("/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506_cleaned4.fam", col.names = c("FID", "IID", "X", "X2", "Sex", "Phenotype"))
fam$Sex <- bulik$reported_gender[match(fam$IID, bulik$chip_well_barcode)]
fam$Sex <- as.vector(ifelse(fam$Sex == "F", 2, ifelse(fam$Sex == "M", 1, 0)))
write.table(fam, "/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506_cleaned_wsex.fam", col.names = F, row.names = F, sep = "\t", quote=F)
```

```{bash}
cp PGCED_SE_ANGI_Bulik_20190506_cleaned4.bed PGCED_SE_ANGI_Bulik_20190506_cleaned_wsex.bed
cp PGCED_SE_ANGI_Bulik_20190506_cleaned4.bim PGCED_SE_ANGI_Bulik_20190506_cleaned_wsex.bim
```


# 8. Remove HH SNPS
```{bash}
#!/bin/bash
#SBATCH --job-name=remove-HH
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink --bfile ${DATA}_cleaned_wsex --set-hh-missing --make-bed --out ${DATA}_noHH
```

```{bash}
cat PGCED_SE_ANGI_Bulik_20190506_noHH.hh | awk -v OFS='\t' '{print $3}'  | uniq > unique_HH.txt
```


# 9. sex concordance 
```{bash}
#!/bin/bash
#SBATCH --job-name=sex_concordance
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink --bfile  ${DATA}_noHH  --check-sex --out SexCheck

```

grep -v OK SexCheck.sexcheck

                  FID                   IID       PEDSEX       SNPSEX       STATUS            F
  201035700195_R04C01   201035700195_R04C01            2            1      PROBLEM       0.9986
  201092190143_R04C02   201092190143_R04C02            2            1      PROBLEM       0.9981
  201096150074_R10C01   201096150074_R10C01            2            0      PROBLEM       0.2821
  201142020027_R03C02   201142020027_R03C02            0            2      PROBLEM      0.06694
  201145010056_R04C02   201145010056_R04C02            2            0      PROBLEM       0.3118
  201148270120_R01C02   201148270120_R01C02            0            2      PROBLEM      0.07245
  201148280064_R06C01   201148280064_R06C01            2            0      PROBLEM       0.2043
  201035700126_R01C01   201035700126_R01C01            2            1      PROBLEM       0.9981
  201035700158_R08C02   201035700158_R08C02            2            0      PROBLEM       0.2245


```{r}
fam <- read.table("/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506_noHH.fam")
fam[fam$V1=="201142020027_R03C02", 5] <- 2
fam[fam$V1=="201148270120_R01C02", 5] <- 2
write.table(fam, "/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506_noHH.fam", col.names = F, row.names = F, sep = "\t", quote=F)
```

# Remove the sample mix up
remove.txt

201035700195_R04C01   201035700195_R04C01
201092190143_R04C02   201092190143_R04C02
201035700126_R01C01   201035700126_R01C01
```{bash}
#!/bin/bash
#SBATCH --job-name=remove-sample
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink --bfile  ${DATA}_noHH  --remove remove.txt --make-bed --out ${DATA}_cleaned5
```


# 10. Align the strand
```{bash}
#!/bin/bash
#SBATCH --job-name=flip-strand
#SBATCH --mem=20GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506
SOFTWARE=/nfs/home/aliwal/software

${SOFTWARE}/update_build.sh  ${DATA}_cleaned5  ${SOFTWARE}/GSA-24v1-0_A1-b37.Source.strand   ${DATA}_flipped
```

# 11. updated the ILLUMINA SNP IDs to the IDs in reference genome GRCh37.p5

```{bash}
#!/bin/bash
#SBATCH --job-name=add_rs_ids
#SBATCH --mem=20GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506
SOFTWARE=/nfs/home/aliwal/software

Rscript ${SOFTWARE}/SNP_to_rsID.R  ${DATA}_flipped.bim
mv ${DATA}_flipped.bim ${DATA}_flipped.bim_ori
mv ${DATA}_flipped.bim_rsed_dbSNP144.GRCh37 ${DATA}_flipped.bim
```

## Heterozygosity check!!
The inbreeding coefficient is also the fraction of all of the genes of an animal that are homozygous (two copies of the same allele). A COI of less than 5% is best. 
```{bash}
#!/bin/bash
#SBATCH --job-name=het
#SBATCH --mem=20GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink --bfile ${DATA}_flipped --het
```

```{r}
het <- read.table("plink.het", header = T)
library(ggplot2)
library(dplyr)
het_05 <- het %>% filter(F > 0.05)
num_outliers <- dim(het_05)[1]
num_outliers

png("Bulik_het_F.png", width=15, height=18, units='in', bg="white", res=150, type=c("cairo"))
ggplot(het, aes(x = F)) + geom_histogram()
dev.off()
```



# 12. Check AF to reference

## VECTOR
```{bash}
#!/bin/bash
#SBATCH --job-name=AF
#SBATCH --mem=20GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506
path2out=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506
plink --bfile ${DATA}_flipped \
          --freq \
          --out ${path2out}/freq_Bulik
```
##DELTA
```{bash}
#!/bin/bash
# 
#PBS -S /bin/bash
#PBS -A UQ-IMB
#PNS -N plink_allele_freq
#PBS -l walltime=2:00:00
#PBS -l select=1:ncpus=1:mem=30GB

DATA=/scratch/90days/uqawal15/1000G_phase3
path2software=/scratch/90days/uqawal15/software
${path2software}/plink --bfile ${DATA}/1000G_phase3_20130502_combined_snpsonly.05 \
          --freq \
          --out ${DATA}/freq_1000G
```

```{bash}
library(data.table)
dat = data.frame(fread("/scratch/90days/uqawal15/1000G_phase3/Bulik/freq_Bulik.frq", header = T, showProgress = F))
KG = data.frame(fread("/scratch/90days/uqawal15/1000G_phase3/freq_1000G.frq", header = T, showProgress = F))
comm <- intersect(dat$SNP, KG$SNP)
dat = dat[dat$SNP %in%comm,]
KG = KG[KG$SNP %in%comm,]

# check allele frequency
dat$KG.maf = KG[match(dat$SNP, KG$SNP),"MAF"]
dat$KG.allele =  KG[match(dat$SNP, KG$SNP),"A1"]

dat$KG.ref = NA
dat[which(dat$A1 == dat$KG.allele), ]$KG.ref = dat[which(dat$A1 == dat$KG.allele) , "KG.maf" ]
dat[which(dat$A2 == dat$KG.allele), ]$KG.ref = (1 - dat[which(dat$A2 == dat$KG.allele) , "KG.maf" ])


dat$diff = NA
dat$diff <- dat$MAF - dat$KG.maf

t = 3
SD0.1 = sd(dat[(dat$KG.maf > 0.09 & dat$KG.maf < 0.11),"diff"]) 
SD0.5 = sd(dat[(dat$KG.maf > 0.49 & dat$KG.maf < 0.51),"diff"])
a = t * (SD0.1 - SD0.5) / 0.16
b = t * SD0.5
fun.up1 =function(x)( x + a*(x-0.5)^2 + b )
fun.dn1 =function(x)( x - a*(x-0.5)^2 - b )

## outliers are outside of the red curves.
dat$outlier = ((dat$diff)^2 > (a*((dat$KG.maf - 0.5)^2) + b )^2)
nrow(dat[dat$outlier == T,])
dat2 <- dat[!is.na(dat$KG.ref),]
mismatch <- dim(dat)[1] - dim(dat2)[1]
print(mismatch)
## write.table(raf[raf$outlier ==T,],"outlier.in.round1.txt", quote = F, sep ="\t")

## plot
library(ggplot2)
png("Bulik_1KG_AF.png", width=15, height=15, units='in', bg="white", res=150, type=c("cairo"))
ggplot(dat2, aes(x=KG.ref, y =MAF)) + 
  geom_point(size=0.5) + 
  stat_function(fun = fun.up1, color = "red") + 
  stat_function(fun = fun.dn1, color = "red") + 
  xlim(0,1) + 
  ylim(0,1)  +  
  geom_abline(intercept=0, slope=1, color="blue") + 
  ggtitle(paste0(t , " times of SD"))
dev.off()

## write table of non-outliers and SNPs that match by alleles
dat3 <- dat2[dat2$outlier == F,]
outliers <- dim(dat2)[1] - dim(dat3)[1]
print(outliers)
write.table(dat3,"non-outliers.txt", quote = F, sep ="\t", col.names = F, row.names =F)
```



# 12. Ancestry Analyses
## DELTA2
```{bash}
cat non-outliers.txt | awk '{print $2}' > non-outliers.snps

### generate GRM
pcref='/scratch/90days/uqawal15/1000G_phase3/1000G_phase3_20130502_combined_snpsonly'
refout='/scratch/90days/uqawal15/1000G_phase3/Bulik/1000G_phase3_20130502_combined_snpsonly'

### generate GRM
/scratch/90days/uqawal15/software/gcta_1.93.3beta2/gcta64 --bfile ${pcref}.05  --extract non-outliers.snps --make-grm --out ${refout}.05.non.outliers
## Calculates PCAs
/scratch/90days/uqawal15/software/gcta_1.93.3beta2/gcta64 --grm    ${refout}.05.non.outliers  --pca 3   --out  ${refout}.05.non.outliers.pca3
## Calculates PC loading
/scratch/90days/uqawal15/software/gcta_1.93.3beta2/gcta64 --bfile  ${pcref}.05  --extract  non-outliers.snps  --pc-loading  ${refout}.05.non.outliers.pca3  --out ${refout}.05.non.outliers.pca3.loading
```

## VECTOR
```{bash}
#!/bin/bash
#SBATCH --job-name=PCA
#SBATCH --mem=20GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00
DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506
pcref='/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/1000G_phase3_20130502_combined_snpsonly'

gcta64  --bfile  ${DATA}_flipped  --extract  non-outliers.snps  --project-loading   ${pcref}.05.non.outliers.pca3.loading  3  --out  ${DATA}_cleaned.pca
```

```{bash}
## PCA plotting
mkdir PCA
DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506
pcref='/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/1000G_phase3_20130502_combined_snpsonly'
STR=/nfs/home/aliwal/software/1000GP_Phase3.sample
Rscript  /nfs/home/aliwal/software/PCA_ancestrycalling_meanSD.R  Bulik    ${pcref}.05.non.outliers.pca3   ${DATA}_cleaned.pca.proj.eigenvec    ${STR}

```

# Table of Samples in each Ancestry
```{r}
anc <- read.table("Bulik_ancestry_assigned_with_meanSD_method.txt", header =T)
```
  AFR   AMR   EAS   EUR other   SAS
   76    81    55  1980    24    26


# Only keep non-outlier SNPs

```{bash}
#!/bin/bash
#SBATCH --job-name=remove-outliers
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

plink --bfile  ${DATA}_flipped  --extract non-outliers.snps --make-bed --out ${DATA}_final
```

# Relatedness of Samples

Only keep European samples
```{bash}
cat Bulik_ancestry_assigned_with_meanSD_method.txt | sed -e '1d' | awk -v OFS='\t' '{if ($7 == "EUR") print $1, $2}' > EUR_samples.txt
```

```{bash}
#!/bin/bash
#SBATCH --job-name=eur
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506
eur=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PCA/EUR_samples.txt

plink --bfile ${DATA}_final --keep ${eur} --make-bed --out ${DATA}_EUR
```

```{bash}
#!/bin/bash
#SBATCH --job-name=GRM
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

gcta64 --bfile  ${DATA}_EUR  --make-grm --out ${DATA}_EUR
```


```{bash}
#!/bin/bash
#SBATCH --job-name=relatives_02
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

gcta64 --grm ${DATA}_EUR --grm-singleton 0.2 --out ${DATA}_EUR_relatedness_02

```

```{bash}
#!/bin/bash
#SBATCH --job-name=relatives_005
#SBATCH --mem=2GB
#SBATCH --nodes=1
#SBATCH --time=0:30:00

DATA=/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506/PGCED_SE_ANGI_Bulik_20190506

gcta64 --grm ${DATA}_EUR --grm-singleton 0.05 --out ${DATA}_EUR_relatedness_005

```

```{r}
prefix="PGCED_SE_ANGI_Bulik_20190506_EUR"
ReadGRMBin=function(prefix, AllN=F, size=4){
  sum_i=function(i){
    return(sum(1:i))
  }
  BinFileName=paste(prefix,".grm.bin",sep="")
  NFileName=paste(prefix,".grm.N.bin",sep="")
  IDFileName=paste(prefix,".grm.id",sep="")
  id = read.table(IDFileName)
  n=dim(id)[1]
  BinFile=file(BinFileName, "rb");
  grm=readBin(BinFile, n=n*(n+1)/2, what=numeric(0), size=size)
  NFile=file(NFileName, "rb");
  if(AllN==T){
    N=readBin(NFile, n=n*(n+1)/2, what=numeric(0), size=size)
  }
  else N=readBin(NFile, n=1, what=numeric(0), size=size)
  i=sapply(1:n, sum_i)
  return(list(diag=grm[i], off=grm[-i], id=id, N=N))
}
GRM = ReadGRMBin(prefix)

png("Bulik_EUR_GRM.png", width=15, height=18, units='in', bg="white", res=150, type=c("cairo"))
hist(GRM$off, breaks = seq(-0.2, 1.2, 0.05), labels = T, main="Distribution of GRM lower triangle", xlab="relationship", cex.main=2, cex.lab=2, cex.axis=2)
dev.off()

off <- GRM$off
off2 <- off[off<0.05]
png("Bulik_EUR_GRM_rm05.png", width=15, height=18, units='in', bg="white", res=150, type=c("cairo"))
hist(off2, labels = T,  breaks = seq(-0.2, 0.06, 0.05), main="Distribution of GRM lower triangle < 0.05", xlab="relationship", cex.main=2, cex.lab=2, cex.axis=2)
dev.off()


off3 <- off[off>=0.05]
png("Bulik_EUR_GRM_greater_05.png", width=15, height=18, units='in', bg="white", res=150, type=c("cairo"))
hist(off3, labels = T,  breaks = seq(0.05, 1.2, 0.05), main="Distribution of GRM lower triangle >= 0.05", xlab="relationship", cex.main=2, cex.lab=2, cex.axis=2)
dev.off()
```


# add relatedness values, het value, and ancestry to extra pheno file. 

```{bash}
library(dplyr)
het <- read.table("plink.het", header = T)
het <- het %>% dplyr::select(FID, IID, F)


anc <- read.table(paste0("/nfs/home/aliwal/ANGI/data/PLINK_GENO/PGCED_SE_ANGI_Bulik_20190506", "/PCA/Bulik_ancestry_assigned_with_meanSD_method.txt"), header = T)
anc <- anc %>% select(FID, population)

final <- merge(het, anc, by = "FID")

fam_05 <- read.table("PGCED_SE_ANGI_Bulik_20190506_EUR_relatedness_005.family.txt", header = F)
related_05 <- unique(fam_05$V1, fam_05$V3)
fam_2 <- read.table("PGCED_SE_ANGI_Bulik_20190506_EUR_relatedness_02.family.txt", header = F)
related_2 <- unique(fam_2$V1, fam_2$V3)

final$related_05 <- as.vector(ifelse(final$population =="EUR" & final$FID %in% related_05, 1, 0))
final$related_2 <- as.vector(ifelse(final$population =="EUR" & final$FID %in% related_2, 1, 0))


write.table(final,"final.pheno", quote = F, sep ="\t", col.names = T, row.names =F)
```

# Move Qced files
```{bash}
mkdir final
mv final.pheno final
mv PGCED_SE_ANGI_Bulik_20190506_final* final
mv PGCED_SE_ANGI_Bulik_20190506_noHH.hh final
```










